Simulation Tools for Small Area Estimation: Introducing the \proglang{R}-package \proglang{saeSim}

### 2015-01-26 saeSim.Rnw
saeSim updates von Timo

%% an abstract and keywords
\Abstract{
The demand for reliable regional estimates from sample surveys \red{has grown substantially} over the last decades. Small area estimation provides statistical methods to produce reliable predictions when the sample sizes in regions are too small to apply direct estimators. Model- and design-based simulations are used to gain insights into the quality of the introduced methods. In this article we present a framework which may help to guarantee the reproducibility of simulation studies in articles and during research. The introduced \proglang{R}-package \proglang{saeSim} is adjusted to provide a simulation environment for the special case of small area estimation. The package may allow the prospective researcher during the research process to produce simulation studies with a minimal effort of coding.   
}

@@

\begin{document}
%
%% include your article here, just as usual
%
% R CMD Sweave %.Rnw
% Loading (and installing) necessary packages.
<<echo=FALSE>>=
# loadAndInstallPackage <- function(packageName) {
#   if(!eval(substitute(suppressPackageStartupMessages(require(packageName))))) {
#     install.packages(packageName)
#     eval(substitute(suppressPackageStartupMessages(library(packageName))))
#   } 
# }
# loadAndInstallPackage("saeSim")
# # Keep order in which packages are loaded, there are naming conflicts between dependencies (MASS and dplyr) which are loaded and attached by the package sae:
# loadAndInstallPackage("sae")
# loadAndInstallPackage("dplyr")
# loadAndInstallPackage("reshape2")
# loadAndInstallPackage("ggplot2")
@
%
\section{Introduction}
The demand for reliable small area statistics from sample surveys \red{has grown substantially} over the last decades due to their use in public and private sectors. In this paper we present a framework for simulation studies inside the field of small area estimation. This tool might be useful for the prospective researcher or data analyst to provide reproducible research.

@@

Next to the combination of text and source code, reproducible research aims at the availability of the full academic research, which is the paper combined with the full computational environment like data and source code. However, real data is often very sensitive and governed by strict confidentiality rules. \red{Synthetic data generation mechanisms as discussed in \cite{alfons2011} and \cite{Kol11} can be used to provide safe data which is publicly available to enable the community to reproduce the analysis and results.} \cite{Bur14} interpreted this as an open research philosophy. Such synthetic data sets can be used to test newly proposed statistical methods in a close-to-reality framework. 
%This may provide valuable insights about the quality of the introduced methods in a controlled environment under different scenarios. 
In general, simulation studies in statistics can be divided into two concepts:
%
\begin{itemize}
\item Design-based: The simulation study is based on true or synthetic data of a fixed population. Then, samples are selected repeatedly from the underlying finite population and different estimation methods are applied in each replication. The obtained estimates are compared to the true values of the population, for instance, in terms of relative bias (RB) or relative root mean squared error (RRMSE).
\item Model-based: The simulation study uses data drawn from certain distributions. In each iteration, the population is generated from a model and a sample is selected according to a specific sampling scheme. The sample is used to estimate the quantity of interest and quality measures (like RB and RRMSE) are derived.  
\end{itemize}
Further discussion regarding model- and design-based simulations is available in \cite{Mue03}, \cite{Sal10} or \cite{Alf10}.

\red{\cite{Alf10} provide the \proglang{R}-package \proglang{simFrame} which helps to conduct simulation studies in a reproducible environment. It includes a wide range of features (like data generation, sampling schemes, outlier contamination mechanisms and missing values) to conduct simulation studies. \proglang{simFrame} was originally developed for simulations in the context of survey statistics but is now designed to be as general as possible \citep[cf.][]{Alf10}. Furthermore the package \proglang{simPop} \citep{Alf14} supports the generation of synthetic population data. This can be a suitable environment in scenarios where the reproducibility of results and confidentiality issues both play an important role.}

@@

However, simulation studies in the context of small area estimation \red{are} often presented very briefly. Thus, there is a need to have a suited framework to guarantee the reproducibility of analysis. To the best of our knowledge, there is not any \proglang{R}-package or framework adjusted for the special case of small area estimation which provides a simulation environment.

@@

\section{Case studies}
\label{sec:caseStudy}
We present two applications of \proglang{saeSim}, one model-based simulation in Section \ref{sec:csModel} and a design-based simulation in Section \ref{sec:csDesign}. Before, we introduce some basic functionalities as the pipe operator (\proglang{\%>\%}) needs some explanation. The pipe operator is designed to make otherwise nested expressions more readable as a line can be read from left to right, instead from inside out \citep{bache14}. \red{As a simple example see the following lines which are equivalent with respect to their functionality:}

@@

In this case the \textit{base-component} is a data frame with an id variable named \proglang{idD} and constructed with the function \proglang{base\_id}. Any random number generator in \proglang{R} can be used. However, we have normally distributed variates, for which some predefined functions are available in the package. \red{For the reproducibility of the following results we also set the seed to 1. The seed is not part of a simulation setup in \proglang{saeSim} but needs to be supplied by the researcher.}

@@

Using this data as population, we repeatedly draw samples from it. Then we predict the domain means by using a direct estimator and a unit-level model. The sampling design is to draw a 10 per cent sample from each region with simple random sampling. For each region the direct estimator for income and the EBLUP under the BHF model is computed. Although the data offers some more information, we only use \proglang{gender}, \proglang{age} and \proglang{agesq} as covariates. The function \proglang{eblupBHF} from the package \proglang{sae} is an implementation of the BHF estimator. This function expects three data objects and returns \red{the} domain predictions. The three objects are the sampled data, the population means of the auxiliary variables and the population sizes in each domain.

\red{Before we begin to construct the simulation setup, we store these data frames as attributes to the population data. However, extra care is needed because not all functions for manipulating data frames in \proglang{R} will preserve their attributes. Functions in \proglang{saeSim} all support this approach. Yet, in user defined functions this needs to be kept in mind by the author. Adding attributes to the data is a very flexible form of processing data on different aggregation levels and we believe this to be a reasonable trade-off. Of course other options can be applied within the possibilities of the \proglang{R} language.}

@@

\red{In the next step, we define components which add the estimates of interest to the data. Here we compute the direct estimator of the mean income in each domain and the EBLUP under the BHF model. Although this could be done in one step, we separate the two computations to illustrate how to combine several estimations and how to define each component independent of each other. This automatically organises the simulation and each component is arranged using the simulation framework. Hence, we define two functions, one for adding the direct estimates and one for adding the EBLUP.}

@@

The package provides more features than introduced in this article. To mention is the support of outlier contaminated data. At this time only representative outliers (outlying observations in the population) are supported \citep[cf.][]{Cha86}. \red{However, we plan to extend this feature to non-representative outliers (outliers are part of the sample but not the population, e.g., incorrectly recorded values). Furthermore, an interface to random number generators in \proglang{R} is available.} The user can also generate group effects as needed in mixed models. As the response is created by an \proglang{R} expression, any form of non-linearity in the relationship between response and auxiliary variables as well as error components can be modelled.

>>>>

%% an abstract and keywords
\Abstract{
The demand for reliable regional estimates from sample surveys has substantially grown over the last decades. Small area estimation provides statistical methods to produce reliable predictions when the sample sizes in regions are too small to apply direct estimators. Model- and design-based simulations are used to gain insights into the quality of the introduced methods. In this article we present a framework which may help to guarantee the reproducibility of simulation studies in articles and during research. The introduced \proglang{R}-package \proglang{saeSim} is adjusted to provide a simulation environment for the special case of small area estimation. The package may allow the prospective researcher during the research process to produce simulation studies with a minimal effort of coding.   
}

@@

\begin{document}
%
%% include your article here, just as usual
%
% R CMD Sweave %.Rnw
% Loading (and installing) necessary packages.
<<echo=FALSE>>=
# loadAndInstallPackage <- function(packageName) {
#   if(!eval(substitute(suppressPackageStartupMessages(require(packageName))))) {
#     install.packages(packageName)
#     eval(substitute(suppressPackageStartupMessages(library(packageName))))
#   } 
# }
# loadAndInstallPackage("saeSim")
# # Keep order in which packages are loaded, there are naming conflicts between dependencies (MASS and dplyr) which are loaded and attached by the package sae:
# loadAndInstallPackage("sae")
# loadAndInstallPackage("dplyr")
# loadAndInstallPackage("reshape2")
# loadAndInstallPackage("ggplot2")
@
%
\section{Introduction}
The demand for reliable small area statistics from sample surveys has substantially grown over the last decades due to their use in public and private sectors. In this paper we present a framework for simulation studies inside the field of small area estimation. This tool might be useful for the prospective researcher or data analyst to provide reproducible research.

@@

Next to the combination of text and source code, reproducible research aims at the availability of the full academic research, which is the paper combined with the full computational environment like data and source code. However, real data is often very sensitive and governed by strict confidentiality rules. Synthetic data generation mechanisms as discussed in \cite{alfons2011} or \cite{Kol11} can be used to provide safe data which is publicly available to enable the community to reproduce the analysis and results. \cite{Bur14} interpreted this as an open research philosophy. Such synthetic data sets can be used to test newly proposed statistical methods in a close-to-reality framework. 
%This may provide valuable insights about the quality of the introduced methods in a controlled environment under different scenarios. 
In general, simulation studies in statistics can be divided into two concepts:
%
\begin{itemize}
\item Design-based: The simulation study is based on true or synthetic data of a fixed population. Then, samples are selected repeatedly from the underlying finite population and different estimation methods are applied in each replication. The obtained estimates are compared to the true values of the population, for instance, in terms of relative bias (RB) or relative root mean squared error (RRMSE).
\item Model-based: The simulation study uses data drawn from certain distributions. In each iteration, the population is generated from a model and a sample is selected according to a specific sampling scheme. The sample is used to estimate the quantity of interest and quality measures (like RB and RRMSE) are derived.  
\end{itemize}
Further discussion regarding model- and design-based simulations is available in \cite{Mue03}, \cite{Sal10} or \cite{Alf10}.

\cite{Alf10} provide the \proglang{R}-package \proglang{simFrame} which helps to conduct simulation studies in a reproducible environment. It includes a wide range of features (like data generation, sampling schemes, outlier contamination mechanisms and missing values) to conduct simulation studies. \proglang{simFrame} was originally developed for simulations in the context of survey statistics but is now designed to be as general as possible \citep[cf.][]{Alf10}. Furthermore the package \proglang{simPop} \citep{Alf14} supports the generation of synthetic population data. This can be a suitable environment in scenarios where the reproducibility of results and confidentiality issues play an important role.

@@

However, simulation studies in the context of small area estimation are often presented very briefly. Thus, there is a need to have a suited framework to guarantee the reproducibility of analysis. To the best of our knowledge, there is not any \proglang{R}-package or framework adjusted for the special case of small area estimation which provides a simulation environment.

@@

\section{Case studies}
\label{sec:caseStudy}
We present two applications of \proglang{saeSim}, one model-based simulation in Section \ref{sec:csModel} and a design-based simulation in Section \ref{sec:csDesign}. Before, we introduce some basic functionalities as the pipe operator (\proglang{\%>\%}) needs some explanation. The pipe operator is designed to make otherwise nested expressions more readable as a line can be read from left to right, instead from inside out \citep{bache14}. As a simple example see the following lines which are equivalent with respect to their functionality:

@@

In this case the \textit{base-component} is a data frame with an id variable named \proglang{idD} and constructed with the function \proglang{base\_id}. Any random number generator in \proglang{R} can be used. However, we have normally distributed variates, for which some predefined functions are available in the package. For the reproducibility of the following results we also set the seed to 1. The seed is not part of a simulation setup in \proglang{saeSim} but needs to be defined by the researcher.

@@

Using this data as population, we repeatedly draw samples from it. Then we predict the domain means by using a direct estimator and a unit-level model. The sampling design is to draw a 10 per cent sample from each region with simple random sampling. For each region the direct estimator for income and the EBLUP under the BHF model is computed. Although the data offers some more information, we only use \proglang{gender}, \proglang{age} and \proglang{agesq} as covariates. The function \proglang{eblupBHF} from the package \proglang{sae} is an implementation of the BHF estimator. This function expects three data objects and returns the domain predictions. The three objects are the sampled data, the population means of the auxiliary variables and the population sizes in each domain.

Before we begin to construct the simulation setup, we store these data frames as attributes to the population data. This allows us to process on the same time a single data frame and some meta data. It is important to notice that some functions for manipulating data frames in \proglang{R} don't preserve their attributes. The prospective user of the \proglang{saeSim} has to keep this in mind when implementing user defined functions. However, all functions in \proglang{saeSim} are able to preserve their attributes. Other options can be applied within the possibilities of the \proglang{R} language. For instance, one could use closures to make the meta data only available for specific functions. This approach needs advanced functional programming skills and may add an additional level of complexity to the package user. However, from our perspective, adding attributes to the data is be a very flexible form of processing data on different aggregation levels and is used in this package.

@@

In the next step, we define components which add the estimates of interest to the data. Here we compute the direct estimator of the mean income in each domain and the EBLUP under the BHF model. Although this could be done in one step, we separate the two computations to illustrate how to combine several estimations and how to define each component independent of each other. This automatically organises the simulation and each component is arranged using the simulation framework. Hence, we define two functions, one for adding the direct estimates and one for adding the EBLUP.

@@

The package provides more features than introduced in this article. To mention is the support of outlier contaminated data. At this time only representative outliers (outlying observations in the population) are supported \citep[cf.][]{Cha86}. However, we plan to extend this feature to non-representative outliers (outliers are part of the sample but not the population, e.g., incorrectly recorded values). Furthermore, an interface to random number generators in \proglang{R} is available. The user can also generate group effects as needed in mixed models. As the response is created by an \proglang{R} expression, any form of non-linearity in the relationship between response and auxiliary variables as well as error components can be modelled.

### 2015-01-26 saeSim.tex
saeSim updates von Timo

\begin{document}
%
%% include your article here, just as usual
%
% R CMD Sweave %.Rnw
% Loading (and installing) necessary packages.
%
\section{Introduction}
The demand for reliable small area statistics from sample surveys \red{has grown substantially} over the last decades due to their use in public and private sectors. In this paper we present a framework for simulation studies inside the field of small area estimation. This tool might be useful for the prospective researcher or data analyst to provide reproducible research.

>>>>

\begin{document}
%
%% include your article here, just as usual
%
% R CMD Sweave %.Rnw
% Loading (and installing) necessary packages.
%
\section{Introduction}
The demand for reliable small area statistics from sample surveys has substantially grown over the last decades due to their use in public and private sectors. In this paper we present a framework for simulation studies inside the field of small area estimation. This tool might be useful for the prospective researcher or data analyst to provide reproducible research.
